### 2016-10-04 capability-machine.tex
added preservation of limits to preordered cofe def, small clarifications

\declaretheorem[numbered=yes,name=Lemma]{lemma}
\declaretheorem[numbered=yes,name=Definition]{definition}
\declaretheorem[numbered=yes,name=Specification]{specification}

@@

%Comments
\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}

@@

\begin{definition}[Preordered \cofes]
  A preordered \cofe{}\ is a \cofe{}\ equiped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. That is the ordering must satisfy the following
  \begin{itemize}
  \item $\forall x \in X \ldotp x \future x$
  \item $\forall x,y,z \in X \ldotp x \future y \land y \future z \Rightarrow x \future z$
  \end{itemize}
\end{definition}

@@

We need the following constructions to create the preordered \cofe{}, we want to
work use \dominique{this sentence doesn't parse :)}
\begin{definition}[Preordered \cofe{}\ construction: Function]
  Given a set $S$ and \cofe{}\ $\var{HP}$, $S \fun \var{HP}$ is a preordered \cofe{}\ with the ordering
\[
  \begin{gathered}
    f \future g \\
    \text{ iff } \\
    \forall s \in \dom(f) \ldotp f(s) \future g(s)
  \end{gathered}
\]
\end{definition}

@@

\subsection{Well-foundedness lemmas}
\begin{lemma}[Future worlds transitive]
  \begin{align*}
    & \forall W, W', W'' \ldotp \\
    & \quad  W'' \future W' \land W' \future W \Rightarrow W'' \future W
  \end{align*}
\end{lemma}

\begin{lemma}[Heap satisfaction uniformity]
  \begin{align*}
    & \forall n,n',\heap, W \ldotp\\
    & \quad \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
  \end{align*}
\end{lemma}

\begin{lemma}[Heap satisfaction non-expansive]
  \begin{align*}
    & \forall n,\heap, W, W' \ldotp\\
    & \quad \heapSat{n}{W} \land W \nequal W' \Rightarrow \heapSat{n}{W'}
  \end{align*}
\end{lemma}

@@

@@

@@

@@

@@

>>>>

\declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
\declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
\declaretheorem[numbered=yes,name=Specification,qed=$\blacksquare$]{specification}

@@

%Comments
\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
\newcommand\kars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}

@@

\begin{definition}[Preordered \cofes]
  A preordered \cofe{}\ is a \cofe{}\ equiped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy chains $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
  % Preorder:
  % \begin{itemize}
  % \item $\forall x \in X \ldotp x \future x$
  % \item $\forall x,y,z \in X \ldotp x \future y \land y \future z \Rightarrow x \future z$
  % \end{itemize}
\end{definition}

@@

We need the following constructions to create the preordered \cofe{} needed to solve the recursive domain equation.
work use \dominique{this sentence doesn't parse :)}
\begin{definition}[Preordered \cofe{}\ construction: Function]
  Given a set $S$ and \cofe{}\ $\var{HP}$, $S \fun \var{HP}$ is a preordered \cofe{}\ with the ordering
\[
  \begin{gathered}
    f \future g \\
    \text{ iff } \\
    \forall s \in \dom(f) \ldotp f(s) \future g(s)
  \end{gathered}
\]
\end{definition}

@@

\subsection{Well-foundedness lemmas}
\begin{lemma}[Future worlds transitive]
  \begin{align*}
    & \forall W, W', W'' \ldotp \\
    & \quad  W'' \future W' \land W' \future W \Rightarrow W'' \future W
  \end{align*}
\end{lemma}
\begin{proof}
  %TODO
\end{proof}

\begin{lemma}
\label{lem:nequal-and-future-world}
  \begin{align*}
    & \forall W_1,W_2,W_1' \ldotp \\
    & \quad W_1 \nequal W_2 \land W_1' \future W_1 \Rightarrow \\
    & \qquad \exists W_2' \ldotp W_2' \future W_2 \land W_1' \nequal W_2'
  \end{align*}
\end{lemma}
\begin{proof}
  Assume $W_1 \nequal W_2$ and $W_1' \future W_2'$. Define $W_2'$ to have $\dom(W_2')=\dom(W_1')$ where
\[
  W_2' \; r = \left\{
  \begin{aligned}
    & (s_1',\phi_2,H_2) & &\text{for } j \in \dom(W_2) \text{ and } W_2 \; j = (s_2,\phi_2,H_2) \\
    &                  & &\quad \text{ and } W_1' \; j = (s_1',\phi',H_1') \\
    & W_1' \; r        & &\text{otherwise}
  \end{aligned} 
  \right.
\]

First we show $W_1' \nequal W_2'$. The domains are defined to be equal, so that is satsfied. Let $r \in \dom(W_1')$ and show $W_1' \; r \nequal W_2' \; r$. First consider the case where $r \in \dom(W_2)$. Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$. We need to show
\begin{align*}
  s_1' & \nequal s_1' \\
  \phi_1' & \nequal \phi_2 \\
  H_1' & \nequal H_2
\end{align*}
The first one is trivial. For the second one we use that we know $\phi_1 \nequal \phi_2$ and $\phi_1 = \phi_1'$ respectively from the $n$-equality assumption and future world assumption. Similarly, the second one follows from $H_1 \nequal H_2$ and $H_1 = H_1'$ - again respectively from the $n$-equality assumption and the future world assumption. For the case where $r \not\in \dom(W_2)$, we need to show $W_1' \; r \nequal W_1' \; r$ which follows trivially from reflexivity.

Next we show $W_2' \future W_2$. We know that $\dom(W_2') = \dom(W_1')$, $\dom(W_1') \supseteq \dom(W_1)$ (from future world assumption), and $\dom(W_1) = \dom(W_2)$ (from $n$-equality assumption). All this can easily be put together to get $\dom(W_2') \supseteq \dom(W_2)$. Now let $r\in \dom(W_2)$ and assume Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$, then we need to show the following:
\begin{align*}
  &(s_2,s_1') \in \phi_2 \\
  &\phi_2 = \phi_2 \\
  &H_2 = H_2
\end{align*}
Here the two last are trivial. The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notive that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
\end{proof}

\begin{lemma}[Heap satisfaction uniformity]
  \begin{align*}
    & \forall n,n',\heap, W \ldotp\\
    & \quad \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
  \end{align*}
\end{lemma}
\begin{proof}
  %TODO
\end{proof}

\begin{lemma}[Heap satisfaction non-expansive]
  \begin{align*}
    & \forall n,\heap, W, W' \ldotp\\
    & \quad \heapSat{n}{W} \land W \nequal W' \Rightarrow \heapSat{n}{W'}
  \end{align*}
\end{lemma}
\begin{proof}
  %TODO
\end{proof}

@@

\begin{lemma}[Read condition non-expansive in worlds]
  \begin{align*}
    & \forall b,e, n, W, W' \ldotp \\
    & \quad W \nequal W' \Rightarrow \\
    & \qquad \forall k < n \ldotp \readCond{k,W,b,e}\\
    & \qquad \quad readCon{k,W',b,e}
  \end{align*}
\end{lemma}
\begin{proof}
  %TODO
\end{proof}

@@

\begin{lemma}[Execute condition non-expansive in worlds]
\label{lem:exec-cond-ne}
  \begin{align*}
    &\forall W_1, W_2,n,b,e,p \ldotp \\
    &\quad W_1 \nequal W_2 \Rightarrow \\
    &\qquad \forall k \leq n \ldotp \\
    &\qquad \quad \execCond{k,W_1,b,e,p} \\
    &\qquad \qquad\Rightarrow \execCond{k,W_2,b,e,p}
  \end{align*}
\end{lemma}
\begin{proof}
  Assume $W_1 \nequal W_2$. Let $k \leq n$ be given. Assume $\execCond{k,W_1,b,e,p}$. Let $k' < k$, $W_2' \future W_2$, and $a \in [b,e]$ be given. We need to show:
\[
  \npair[k']{(p,b,e,a)} \in \stder(W_2')
\]
By lemma~\ref{lem:nequal-and-future-world} there exists a future world $W_1'$ such that $W_1' \nequal W_2'$ and $W_1' \future W_1$. Now use the execute condition assumption with $k' < k$, $W_1' \future W_1$, and $a$ to get
\[
  \npair[k']{(p,b,e,a)} \in \stder(W_1')
\]
Be lemma~\ref{lem:stder-ne}, we know $\stder$ is non-expansive, so $\stder(W_1') \nequal \stder(W_2')$. As $k' < n$, we got the result we wanted.
\end{proof}

@@

\begin{lemma}[Entry condition non-expansive in worlds]
  \begin{align*}
    &\forall W_1, W_2,n,b,e,a \ldotp \\
    &\quad W_1 \nequal W_2 \Rightarrow \\
    &\qquad \forall k \leq n \ldotp \\
    &\qquad \quad \entryCond{k,W_1,b,e,a} \\
    &\qquad \qquad\Rightarrow \entryCond{k,W_2,b,e,a}
  \end{align*}
\end{lemma}
\begin{proof}
  Similar to the proof of Lemma~\ref{lem:exec-cond-ne}.
\end{proof}

@@

\begin{lemma}[Value relation contractive]
  \begin{align*}
    & \forall n \ldotp \forall W \in \Worlds \forall w \ldotp \\
    & \quad \npair[n]{w} \in \stdvr(W) \Rightarrow \\
    & \qquad \npair[n+1]{w} \in \stdvr(W)
  \end{align*}
\end{lemma}
\begin{proof}
  %TODO
\end{proof}

@@

\begin{lemma}[Expression relation non-expansive in world]
\label{lem:stder-ne}
\begin{align*}
  \forall W_1 \nequal W_2 \ldotp \stder(W_1) \nequal \stder(W_2)
\end{align*}
\end{lemma}
